Waving and swaying

### 2014-08-15 Paper/grass.tex
Aug15

\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress,10pt]{revtex4-1}  % for review and submission
%\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
%\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
%\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
%\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{slashbox}
\usepackage{color}
\usepackage{longtable}
\usepackage{array}
\usepackage{dashrule}
\usepackage{dcolumn}% Align table columns on decimal point
%\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{ifthen}
\usepackage{amsthm} % Theorem Formatting
\usepackage{amssymb}	% Math symbols such as \mathbb
\usepackage{calrsfs}
\usepackage{subfigure}
\graphicspath{ {images/} }
\newcommand{\ellcap}{{\ell_\text{c}}}
\newcommand{\Fh}{{\boldsymbol{F}_h}}
\newcommand{\Fv}{{\boldsymbol{F}_v}}
\newcommand{\Tv}{{\boldsymbol{T}}}
\newcommand{\bn}{{\boldsymbol{n}}}
\newcommand{\bt}{{\boldsymbol{t}}}
\newcommand{\bz}{{\boldsymbol{z}}}
\newcommand{\bx}{{\boldsymbol{x}}}
\newcommand{\by}{{\boldsymbol{y}}}
\newcommand{\bT}{{\boldsymbol{T}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}

@@

\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87,Nepf99}. 
The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper. 
\newline
Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies.
Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 
\newline   
Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
%Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
%are exhibited in form of large amplitude oscillations knows as mo-nami.
\newline
While shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations,
several aspect of existing theory remain unsatisfactory. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical 
Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$. Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability. These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of phenomenon.

@@

continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance \shreyas{Where did this model come from? Has such a model been used before? Citations to those needed.} as

@@

$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $??<y<<??$ and $S(y)=0$ in the unvegetated layer $?? < y<< ??$. The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=??$) and top ($y=??$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{fig:basicflow}. Flow consist of a region of approximately constant velocity

@@

  \includegraphics[scale=0.8]{fig1}
\caption{Basic flow profile (left )and model of canopy (bottom) along with variation of shear layer (inset) }

@@

\includegraphics[]{Critical_Re_vs_delta}
\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
\label{Re_vs_delta}
Solution of \eqref{Orr-somerfield} predicts a critical $R$ as function of grass number density, below which the flow is stable (Re($\sigma$)$<0$) and above which it is unstable (Re($\sigma$)$>0$),
as shown in Figure\ref{Re_vs_delta}. The solution also predicts a finite critical wavenumber $k$ corresponding to this threshold for instability shown in Figure \ref{K_vs_shear}. 
\shreyas{Needs language clean up from here on. Describe the most important observation. The details are to be explained in the longer paper.}
we can observe that above a 
certain threshold $\delta$ the critical wavelength decreases with the decrease in $\delta$, similar to the one predicted by K-H theory based on the instability of free shear flow, for that
reason this region is identified as shear layer region. We also observe that once the $\delta$ is decreased below the threshold value, there isn't any variation in the 
critical wavelength with the decrease of shear layer thickness and $\lambda_{critical} \sim (H-h_{grass})$, we identify this region as shear instability of unvegetated region.
The boundary of transition from instability of shear-layer near the tip of grass to the instability of unvegetated flow depends on the submergence ratio of grass. 
\includegraphics[]{K_vs_shear_width_noshear}
\caption{Critical Wavenumber for different grass height as function of shear layer thickness}
\label{K_vs_shear}
% \begin{figure}[htb]
% \includegraphics[]{Low_drag_k}
% \caption{Critical Wavenumber for different grass height as function of $f(\delta)$}
% \label{Low_drag_k}
% \end{figure}

@@

The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness $5 <H/\delta <800$ for canopy with low submergence ratio $h/H\leq 0.4$.
Canopy with low submergence ratio can also be related to the canopy in unbounded flow, such as terrestrial canopy with atmospheric flow with $h/H\sim 0$. 
The linear variation of critical wavelength with the shear layer thickness in this regime is also consistent with the observed wavelength of honami for terrestrial canopies.
However, the deviation of critical wavelength from the prediction based on the instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 
regimes where flow instability does not arise from mechanism similar to Kelvin-Helmholtz instability. The instability in region-1 is understood to be arising from instability of flow
comprising flow in shear layer and flow above the canopies. The instability of flow comprising flow in shear layer and flow above the vegetation can be endorse by the linear 
variation of the critical wavelength with the $H-h+b\delta$ in Figure \ref{Low_drag_k}, where $b\delta$ is the amount of perturbation of the flow into the vegetation.

@@

The constant value of critical wavelength for all the shear layer smaller than a critical value in region-3 suggest that the instability 
in the region-3 is arising from the shear instability of flow above the canopy. The existence of a regime dominated by shear instability of flow above grass is further corroborated 
by the collapse of eigen-modes for different $\delta<\delta_{cr}$ ($\delta/H=0.0037,0.0034,0.0024$ in Figure\ref{Asymptotic_mode}) into a single profile as shown in 
Figure \ref{Asymptotic_mode}, where $\delta_{cr}$ is the critical shear layer thickness
below which instability of vegetated flow arise through the mechanism of shear instability of flow in unvegetated region rather than through the mechanism of Kelvin-Helmholtz type 
instability of shear layer ($\delta/H=0.0052,0.0042$ in Figure \ref{Asymptotic_mode}). The regime of shear instability of flow above the 

@@

Numerical investigation of flow in presence of canopy reveal three distinct zone of flow instability which might lead to monami. The occurrence of these three flow instability in a canopy
depends on its submergence ratio and shear layer thickness. Our analysis shows that flow in presence of grass become unstable not only through the mechanism similar to Kelvin-Helmholtz
instability but also through shear instability of flow above the grass for small shear layer thickness and shear instability of flow comprising flow in shear layer with flow above the grass.
Our analysis further predict a threshold criteria characterized by a Reynolds number above which flow become unstable. The predicted value of threshold Reynolds number is $O(100)$, which is 
close to the the Reynolds number corresponding to the observed value of threshold flow speed in the observation of Grizzle. Inclusion of variation of drag coefficient along the grass filament

@@

%\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}
%\cite{Nepf02}\cite{Nepf04}\cite{Nepf08}\cite{Nepf08_2}\cite{Raupach11}\cite{Raupach94}
%\cite{Delangre06}
%\begin{figure}[htb]Q
%\subfigure{\includegraphics[]{Vivoni_Fig3_6_zero_shear_match}}
%\subfigure{\includegraphics[]{Vivoni_Fig3_7_zero_shear_match}}
%\caption{Comparison between observed and calculated velocity profile}
%\end{figure}
%\begin{figure}[htb]
%\includegraphics[]{Vivoni_Fig3_7_hb_14match}
%\caption{Comparison between observed and calculated velocity profile,hb=14}
%\end{figure}
%\begin{figure}[htb]
%\includegraphics[]{Vivoni_Fig3_7_hb_17match}
%\caption{Comparison between observed and calculated velocity profile,hb=17}
%\end{figure}
\bibliography{Grass}{}
\bibliographystyle{plain}
%\nocite{*}
\end{document}

>>>>

\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress,10pt]{revtex4-1}  % for review and submission
%\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
%\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
%\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
%\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{slashbox}
\usepackage{color}
\usepackage{longtable}
\usepackage{array}
\usepackage{dashrule}
\usepackage{dcolumn}% Align table columns on decimal point
%\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{ifthen}
\usepackage{amsthm} % Theorem Formatting
\usepackage{amssymb}	% Math symbols such as \mathbb
\usepackage{calrsfs}
\usepackage{subfigure}
\graphicspath{ {images/}{images/Growth_Rate_Cont_Plots/} }
\newcommand{\ellcap}{{\ell_\text{c}}}
\newcommand{\Fh}{{\boldsymbol{F}_h}}
\newcommand{\Fv}{{\boldsymbol{F}_v}}
\newcommand{\Tv}{{\boldsymbol{T}}}
\newcommand{\bn}{{\boldsymbol{n}}}
\newcommand{\bt}{{\boldsymbol{t}}}
\newcommand{\bz}{{\boldsymbol{z}}}
\newcommand{\bx}{{\boldsymbol{x}}}
\newcommand{\by}{{\boldsymbol{y}}}
\newcommand{\bT}{{\boldsymbol{T}}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\grad}{\mathbf{\nabla}}
\newcommand{\del}{\partial}

@@

%\section{Introduction}
Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
Resulting changes in hydrodynamic conditions can influence number of environmental variables and processes such as 
transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87,Nepf99}. 
The response of flexible grass canopies to steady current in form of large amplitude coherent oscillation, known as mo-nami \cite{AckermanOkubo93}, plays a central role
in the recruitment of microscopic marine organism such as blue mussel larvae \cite{Grizzle96}. The hydrodynamic mechanism underlying monami is the focus of this paper. 
\newline
Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.
A crucial difference between the atmospheric and aquatic flow is that the atmospheric flows are essentially unbounded vertically. Another major
difference between the two is the considerable difference of stiffness of canopies; terrestrial canopies tend to be much more rigid than aquatic canopies.
Due to these differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 
\newline   
Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to
different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.
%Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
%are exhibited in form of large amplitude oscillations knows as mo-nami.
\newline
While shear layer model successfully appears to predict frequency of mo-nami for number of experimental observations,
several aspect of existing theory remain unsatisfactory. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $, whereas observed critical 
Reynolds number\cite{Grizzle96} (Grizzle 1996) for monami is much higher $\approx O(1000)$. Finally, the thickness of the free shear layer, defined as the momentum thickness of the boundary layer near the grass tip, is in many cases comparable to the unvegetated layer thickness, and therefore inconsistent with the assumptions of the free shear layer instability. These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of phenomenon.

@@

continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} \shreyas{Where did this model come from? Has such a model been used before? Citations to those needed.} as

@@

$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $0<y<h_{grass}$ and $S(y)=0$ in the unvegetated layer $h_{grass}< y< H$. 
The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=0$) and top ($y=H$) of the fluid layer, as can be seen from the comparison 
between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{fig:basicflow}. Flow consist of a region of approximately constant velocity

@@

  \includegraphics[scale=1]{Grass_Base_Vivoni_shear}
\caption{Basic flow profile (left )from the experiment and model and the model of canopy (bottom) along with variation of shear layer (inset) }

@@

\begin{figure}[htb]
\includegraphics[]{Critical_Re_vs_delta}
\caption{Critical Reynolds number with flow for different grass height as a function of inverse shear thickness}
\label{Re_vs_delta}
\end{figure}
Solution of \eqref{Orr-somerfield} predicts a critical $R$ as function of grass number density, below which the flow is stable (Re($\sigma$)$<0$) and above which it is unstable (Re($\sigma$)$>0$),
as shown in Figure\ref{Re_vs_delta}. The solution also predicts a finite critical wavenumber $k$ corresponding to this threshold for instability shown in Figure \ref{K_vs_shear}. 
\shreyas{Needs language clean up from here on. Describe the most important observation. The details are to be explained in the longer paper.}
We observe that above a 
certain threshold $\delta$ the critical wavenumber is inversely proportional to $\delta$, similar to the one predicted by K-H theory based on the instability of free shear flow,
we identify this region as shear layer region. We also observe that once the $\delta$ is decreased below a critical value, there isn't any variation in the 
critical wavelength with the decrease of shear layer thickness and $\lambda_{critical} \sim (H-h_{grass})$, we identify this region as shear instability of unvegetated region.
The boundary of transition from instability of shear-layer near the tip of grass to the instability of unvegetated flow depends on the submergence ratio of grass. 
\begin{figure}[htb]
\includegraphics[]{K_vs_shear_width_noshear}
\caption{Critical Wavenumber for different grass height as function of shear layer thickness}
\label{K_vs_shear}
\end{figure}
\begin{figure*}
%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens31_imgsc}}
%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens32_imgsc}}
%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens34_imgsc}}
%\subfigure[]{\includegraphics[scale = 0.24]{Set3_dens38_imgsc}}

\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens28_imgsc}}
\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens30_imgsc}}
\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens32_imgsc}}
\subfigure[]{\includegraphics[scale = 0.55]{Set4_dens34_imgsc}}

\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens38_imgsc}}
\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens40_imgsc}}
\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens42_imgsc}}
\subfigure[]{\includegraphics[scale = 0.55]{Set5_dens46_imgsc}}

\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens32_imgsc}}
\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens34_imgsc}}
\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens36_imgsc}}
\subfigure[]{\includegraphics[scale = 0.55]{Set6_dens38_imgsc}}

\caption{$Re(\sigma)$ as function of wavenumber and Reynolds number for a given grass number density along with contour of zero growth rate ($Re(\sigma)$=0). The figures in each 
row corresponds to a particular grass height $h_{grass}/H = 0.8$ for (a-d), 1.0 for (e-h) and  1.2 for (i-l), these figures are arranged in order of increasing grass 
number density. At low grass number density both the modes are indistinguishable, but as $N_g$ increases the distinction between the two modes become clear. For $N_g$ below
a critical value, the K-H mode sets the onset of flow instability whereas above the critical $N_{g}$ onset of flow instability is dominated by shear instability of flow above the grass}
\label{K_Re_sigma_set3}
\end{figure*}

@@

The transition of dominant mode of flow instability from K-H mode to shear instability of flow above the grass can be further corroborated from Figure \ref{K_Re_sigma_set3}.
We observe that the two different modes of instability are indistinguishable at low grass density. The distinction becomes apparent with the increase in grass number density, where we 
can see that dominant mode of flow instability is 1st dominated by K-H mode before switching to shear instability of flow above the grass at relatively large grass density.  
%\newline
%The mechanism similar to Kelvin-Helmholtz instability is observed for a wide range of shear layer thickness for canopy with low submergence ratio $h/H\leq 0.4$.
%Canopy with low submergence ratio can also be related to the canopy in unbounded flow, such as terrestrial canopy with atmospheric flow with $h/H\sim 0$. 
%The linear variation of critical wavelength with the shear layer thickness for terrestrial canopy ($h/H \sim 0$) is consistent with the observed wavelength of honami 
%for terrestrial canopies.
%However, the deviation of critical wavelength from the prediction based on the instability of shear layer in region-1 and region-3 for a given submergence ratio shows the existence of 
%regimes where flow instability does not arise from mechanism similar to Kelvin-Helmholtz instability. The instability in region-1 is understood to be arising from instability of flow
%comprising flow in shear layer and flow above the canopies. The instability of flow comprising flow in shear layer and flow above the vegetation can be endorse by the linear 
%variation of the critical wavelength with the $H-h+b\delta$ in Figure \ref{Low_drag_k}, where $b\delta$ is the amount of perturbation of the flow into the vegetation.

@@

%The constant value of critical wavelength for all the shear layer smaller than a critical value in region-3 suggest that the instability 
%in the region-3 is arising from the shear instability of flow above the canopy. 
%The existence of a regime dominated by shear instability of flow above grass is further corroborated 
%by the collapse of eigen-modes for different $\delta<\delta_{cr}$ ($\delta/H=0.0037,0.0034,0.0024$ in Figure\ref{Asymptotic_mode}) into a single profile as shown in 
%Figure \ref{Asymptotic_mode}, where $\delta_{cr}$ is the critical shear layer thickness
%below which instability of vegetated flow arise through the mechanism of shear instability of flow in unvegetated region rather than through the mechanism of Kelvin-Helmholtz type 
%instability of shear layer ($\delta/H=0.0052,0.0042$ in Figure \ref{Asymptotic_mode}). 
The regime of shear instability of flow above the 

@@

Numerical investigation of flow in presence of canopy reveal two distinct modes of flow instability which might lead to monami. The occurrence of these modes for flow instability in a 
canopy depends on its submergence ratio and shear layer thickness. Our analysis shows that flow in presence of grass become unstable not only through the mechanism similar to 
Kelvin-Helmholtz instability but also through shear instability of flow above the grass for small shear layer thickness and shear instability of flow comprising flow in shear layer 
with flow above the grass. Our analysis further predict a threshold criteria characterized by a Reynolds number above which flow become unstable. The predicted value of threshold
 Reynolds number is $O(100)$, which is 
close to the the Reynolds number corresponding to the observed value of threshold flow speed in the observation of Grizzle. Inclusion of variation of drag coefficient along the 
grass filament

@@

%\cite{Delangre06}\cite{Delangre04}\cite{Raupach96}\cite{Nepf99}
%\cite{Nepf02}\cite{Nepf04}\cite{Nepf08}\cite{Nepf08_2}\cite{Raupach11}\cite{Raupach94}
%\cite{Delangre06}
%\begin{figure}[htb]Q
%\subfigure{\includegraphics[]{Vivoni_Fig3_6_zero_shear_match}}
%\subfigure{\includegraphics[]{Vivoni_Fig3_7_zero_shear_match}}
%\caption{Comparison between observed and calculated velocity profile}
%\end{figure}
%\begin{figure}[htb]
%\includegraphics[]{Vivoni_Fig3_7_hb_14match}
%\caption{Comparison between observed and calculated velocity profile,hb=14}
%\end{figure}
%\begin{figure}[htb]
%\includegraphics[]{Vivoni_Fig3_7_hb_17match}
%\caption{Comparison between observed and calculated velocity profile,hb=17}
%\end{figure}
\bibliography{Grass}{}
\bibliographystyle{plain}
%\nocite{*}
\end{document}
